# 1. Read Data

```{r}
data(longley)
y <- as.matrix(longley$Employed)
df <- longley[, -which(names(longley) == "Employed")]
X <- as.matrix(cbind(1, df))  # Add intercept (ones) as the first column
X
```

# 2. Compute Models and Time Consumed

We find that LU decomposition is the fastest method comparing to others, Cholesky decomposition method is the second.

#### LU Decomposition

```{r}
start_time <- Sys.time()
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y  # Compute beta_hat
residuals <- y - X %*% beta_hat
n <- nrow(X)
p <- ncol(X)
sigma_hat_sq <- t(residuals) %*% residuals / (n - p)  # Compute sigma_hat_sq
var_beta_hat <- as.numeric(sigma_hat_sq) * solve(t(X) %*% X)  # Compute variance of beta_hat
se_beta_hat <- sqrt(diag(var_beta_hat))  # Compute SE of beta_hat
end_time <- Sys.time()
time_GE_LU <- end_time - start_time
time_GE_LU
```

#### Cholesky decomposition

```{r}
start_time <- Sys.time()
XtX <- t(X) %*% X
XtX_inv <- chol2inv(chol(XtX))
beta_hat_chol <- XtX_inv %*% t(X) %*% y
residuals_chol <- y - X %*% beta_hat_chol
sigma_hat_sq_chol <- t(residuals_chol) %*% residuals_chol / (n - p)
var_beta_hat_chol <- as.numeric(sigma_hat_sq_chol) * XtX_inv
se_beta_hat_chol <- sqrt(diag(var_beta_hat_chol)) 
end_time <- Sys.time()
time_Cholesky <- end_time - start_time
time_Cholesky
```

#### QR decomposition

```{r}
start_time <- Sys.time()
qr_decomp <- qr(X)  # QR decomposition of X
R_inv <- solve(qr.R(qr_decomp))  # Inverse of R
Q <- qr.Q(qr_decomp)  # Q matrix

beta_hat_qr <- R_inv %*% t(Q) %*% y  # Compute beta_hat

residuals_qr <- y - X %*% beta_hat_qr
sigma_hat_sq_qr <- t(residuals_qr) %*% residuals_qr / (n - p)  # Compute sigma_hat_sq

var_beta_hat_qr <- as.numeric(sigma_hat_sq_qr) * R_inv %*% t(R_inv)  # Compute variance of beta_hat
se_beta_hat_qr <- sqrt(diag(var_beta_hat_qr))  # Compute SE of beta_hat
end_time <- Sys.time()
time_QR <- end_time - start_time
time_QR
```

#### Check the results

```{r}
model <- lm(Employed ~ ., data = longley)
summary(model)
beta_hat
beta_hat_chol
beta_hat_qr
```

# 3. Prove The Ridge Solotion

$J(\beta) = (1/2)||y-X\beta||^2 + (λ/2)||\beta||^2$

$\frac {\partial J(B)}{\partial \beta} = -X^T(y-X \beta) + \lambda \beta = 0$

$-X^Ty+X^TX \beta + \lambda \beta = 0$

$(X^TX + \lambda I) \beta = X^T y$

$\hat{\beta} = (X^TX + \lambda I)^{-1} X ^Ty$

# 4. Plot 2-norm of Ridge Function

```{r}
lambda_values <- seq(0, 100, by = 50)
norms <- numeric(length(lambda_values))
for (i in seq_along(lambda_values)) {
  lambda <- lambda_values[i]
  beta_hat_ridge <- solve(t(X) %*% X + lambda * diag(p), t(X) %*% y)
  norms[i] <- sqrt(sum(beta_hat_ridge^2))
}
plot(lambda_values, norms, type = "l", xlab = "Lambda", ylab = "L2-Norm of Ridge Coefficients", ylim = c(0,8000))

```

# Using Parallel Computing

```{r}
# load the parallel package
library(parallel)

lambda_values <- seq(0, 100, by = 50)
p <- ncol(X)

# create a function to calculate beta_hat_ridge and its norm
calc_norm <- function(lambda) {
  beta_hat_ridge <- solve(t(X) %*% X + lambda * diag(p), t(X) %*% y)
  norm <- sqrt(sum(beta_hat_ridge^2))
  return(norm)
}

# use mclapply to calculate norms in parallel
norms <- mclapply(lambda_values, calc_norm, mc.cores = detectCores())

# since mclapply returns a list, unlist norms to a numeric vector
norms <- unlist(norms)

# plot the results
plot(lambda_values, norms, type = "l", xlab = "Lambda", ylab = "L2-Norm of Ridge Coefficients", ylim = c(0,8000))

```

# Find out which method is the lm() function in R is using? And which algorithm is being used?

QR decomposition is the method that lm() function in R is using.

QR algorithm factor a matrix into a product of an orthogonal matrix Q and a upper triangular matrix R, and use them to solve the linear equation,

$X \beta = y$

$R \beta = Q^Ty$

, by solving the second equation, we found that it's easier to compute than the first one.

But it's not always this case, R actually will choose the method that will depends on the circumstance. Sometimes, SVD method will be used to approach numeric stability, but it's more computationally expensive.
